Complete Algebraic File Synchronization

### 2015-12-23 2015/axioms.tex
Comparison proof and text improvements

So that we could describe the effects of commands on a filesystem, let us introduce some notation
and note some observations.
We already know that
commands usually do not occur in isolation,
and are applied to filesystems in time.
Therefore we will investigate sequences of commands with a well-defined order
(which form a free semigroup).
We use $\cc$ to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbdaa{n}{\vald}\cc\cbfaa{m}{\valf}$, or $S'=S\cc\cbba{o}$.
We write $\emptyseq$ for the empty sequence.

@@

We methodically investigated of all possibilities using a computer 
program\footnote{The program is accessible on-line at \\
https://github.com/csirmaz/AlgebraicSyncPaper/blob/master/2015/prove.py.}
to determine
which pairs of commands cause errors all the time,
which can be simplified to one or no commands, and which can be reversed
without any change in their overall effect.
Our basis for this investigation was the model of filesystems introduced in this paper.
These properties of command pairs are crucial as they determine
how a set of commands can be re-ordered to be applied to a filesystem
during synchronization, and what commands will never be compatible.

Below we list a number of statements derived using this method.
Although they can be proven by investigating all relevant possibilities
in a filesystem, we refer to these as axioms because in a secondary algebraic model
that describes commands without reference to an underlying filesystem,
these would function as meta-rules generating statements we would accept as true.
See \cref{sec_algebra} for more about such a model.

\medskip

Pairs of commands in general have the form
\[ \cxynv\cc  \czwmv \]
where $X,Y,Z,W\in\typeset$, $n,m\in\setn$, 
and values are of the appropriate type: 
$\valvy\in\setvx{Y}$ and $\valvw\in\setvx{W}$.

@@

\begin{mydef}{Construction pair.}
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{gather*}
            \cbdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
            \cbdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii} \\ % TODO all directories are the same?
            \cfdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
            \cfdaa{\nn}{\valdi}\cc  \cbdaa{n}{\valdii} % TODO all directories are the same?
   \end{gather*}
\end{mydef}

@@

\begin{mydef}{Assertion command.}
A command is an assertion command if
for every filesystem it either breaks it or leaves it in the same state.
In other words, assertion commands have the same input and output type,
and there must also be only one possible value for their output.
Accordingly, in our current algebra, $\cbba{n}$ and $\cdda{n}$ are the only two assertion commands.
\end{mydef}

\begin{myax}\label{ax_child_assert}
An assertion command can be added on a descendant node:
$\cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n}$ 
where $\nn\descendant n$ and $\cxy\neq\cdd$.
\end{myax}

>>>>

So that we could describe the effects of commands on a filesystem, let us introduce some notation
and note some observations.
We already know that
commands usually do not occur in isolation,
and are applied to filesystems in time.
Therefore we will investigate sequences of commands with a well-defined order
(which form a free semigroup).
We use ``$\cc$'' to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbdaa{n}{\vald}\cc\cbfaa{m}{\valf}$, or $S'=S\cc\cbba{o}$.
We write $\emptyseq$ for the empty sequence.

@@

We methodically investigated of all possibilities using a computer 
program\footnote{The program is accessible on-line at \\
https://github.com/csirmaz/AlgebraicSyncPaper/blob/master/2015/prove.py.}
to determine
which pairs of commands cause errors all the time,
which can be simplified to one or no commands, and which can be reversed
without any change in their overall effect.
Our basis for this investigation was the model of filesystems introduced in this paper.
These properties of command pairs are crucial as they determine
how a set of commands can be re-ordered to be applied to a filesystem
during synchronization, and what commands will never be compatible.
Below we list a number of statements derived using this method.

\bigskip

\noindent
Pairs of commands in general have the form
\[ \cxynv\cc  \czwmv \]
where $X,Y,Z,W\in\typeset$, $n,m\in\setn$, 
and values are of the appropriate type: 
$\valvy\in\setvx{Y}$ and $\valvw\in\setvx{W}$.

@@

\begin{mydef}{Construction pair.}
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{gather*}
            \cbdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
            \cbdaa{\nn}{\vald}\cc  \cbdaa{n}{\vald} \\
            \cfdaa{\nn}{\vald}\cc  \cbfaa{n}{\valf} \\
            \cfdaa{\nn}{\vald}\cc  \cbdaa{n}{\vald}
   \end{gather*}
\end{mydef}

@@

\begin{mydef}{Assertion command.}
A command is an assertion command if
for every filesystem it either breaks it or leaves it in the same state.
In other words, assertion commands have the same input and output type,
and there must also be only one possible value for their output.
Accordingly, $\cbba{n}$ and $\cdda{n}$ are the only two assertion commands.
\end{mydef}

\begin{myax}\label{ax_child_assert}
An assertion command can be added on a descendant node:
$\cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n}$ 
where $\nn\descendant n$ and $\cxy\neq\cdd$.
(Note that this and next rule is true because the assertion command is
adjacent to the original command.)
\end{myax}

### 2015-12-23 2015/comparisons.tex
Comparison proof and text improvements

\subsection{Liberal and conservative Reconciliation}
We consider the reconciliation algorithm described here an improvement over
the one derved during our previous research (see \cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

@@

>>>>

\subsection{Liberal and conservative Reconciliation}

We consider the reconciliation algorithm described here an improvement over
the one derved during our previous research (see \cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

@@

\subsection{Comparing Definitions of Conflicts}

As a test of the proposed reconciliation algorithm, 
we compare our definition of conflicting
updates to how conflicts are defined by Balasubramaniam and Pierce
in their state-based approach (\cite{BP}).
The update detector described in \cite{BP} provides a safe estimate of nodes
(paths) at which updates occurred.
It marks some nodes as \emph{dirty} in a way that we know that at non-dirty nodes
the filesystem have not changed between their original and current states
(see Definition 3.1.1 in \cite{BP}).
The \emph{dirty} marks are also up-closed, that is, all ancestor nodes of a dirty node
are also dirty (Fact 3.1.3 in \cite{BP}).
And finally, a conflict is detected between replicas $\FS_A$ and $\FS_B$ at node $n$
if $n$ is marked as dirty in both $\FS_A$ and $\FS_B$, and
$\FS_A(n)\neq\FS_B(n)$, and $n$ does not point to a directory in both replicas
($\FS_A(n)\not\in\setd \vee \FS_B(n)\not\in\setd$; see section 4.1 in \cite{BP}).

It can be easily seen that not all conflicts detected based on the above definition
entails a conflict based on our system; that is, \cite{BP} describes a more
conservative policy.
For example, if $\nn=\parent(n)$ and $\FS(\nn)=\vald$ and $\FS(n)=\valf$, and
the two replicas are derived in the following way:
\begin{gather*}
\FS_A = \cfba{n} \cc \cdba{\nn} \FS \\
\FS_B = \cfba{n} \FS,
\end{gather*}
then $\nn$ is dirty in both replicas:
in $\FS_A$ it was modified, and in $\FS_B$ one of its descendants was modified.
Moreover, $\FS_A(\nn)\neq\FS_B(\nn)$ and $\FS_A(\nn)=\empt\not\in\setd$,
and therefore a conflict is detected at $\nn$.
Our reconciliation algorithm detects no conflicts;
instead, it propagates $\cdba{\nn}$ to $\FS_B$, which we think is as expected
and desired.

At the same time, it can be shown that if our command-based reconciler
detects a conflict, it entails a conflict in the state-based reconciler.
We note here that Balasubramaniam and Pierce also suppose that all directories are equal,
therefore, as elsewhere, we are safe to continue to assume that $|\setd|=1$.

\begin{proof}
Let $A$ and $B$ be two simple sequences returned by the update detector
for the two replicas $\FS_A$ and $\FS_B$.
A conflict between the commands $\cxynv\in A$ and $\czwmv\in B$
means that even though $\cxynv\in A\setminus B$
and $\czwmv\in B\setminus A$, they cannot be included in
$\reca$ and $\recb$, respectively, because
$\cxynv\nindep\czwmv$ (see \cref{def:reconciliation}).

From \cref{incomparable_is_independent} we therefore know that $n\nunrel m$.
Without loss of generality we can assume that $n\descendantEq m$.
From this we see that $n$ is \emph{dirty} in both $\FS_A$ and $\FS_B$,
as the filesystem changes at $n$ in $\FS_A$ and at $n$ or its descendant in $\FS_B$
(and none of the commands are assertion commands).

Now we only need to show that $\FS_A(n)\neq\FS_B(n)$, because from this 
and $|\setd|=1$ we will also
know that at least one of these values is not a directory,
and therefore there is a conflict at $n$ in the state-based reconciler.
If $n=m$, this follows from the fact that $\cxynv\neq\czwmv$ 
(because they are not in $A\cap B$),
as from \cref{worksinputmatch} we have $X=Z$,
and therefore the above can only be true
if $\FS_A(n)=\valvy\neq\valvw=\FS_B(n)$.
If $n\descendant m$ and there is no command on $n$ in $B$, then it is also
true as $\cxynv$ is not an assertion command, and therefore 
$\FS_A(n)\neq\FS(n)=\FS_B(n)$, where $\FS$ is the common ancestor of the replicas.
Finally, if there is a command on $n$ in $B$, then 
from $\cxynv\in A\setminus B$ we know
it must be different from $\cxynv$, and similarly to the first case,
we have $\FS_A(n)\neq\FS_B(n)$.
\end{proof}

### 2015-12-23 2015/def-commands.tex
Comparison proof and text improvements

Next we define commands on the filesystem.
As described above, we aimed to select a set of commands
that capture as much information as possible
about the operations, and is also symmetric.
We start by summarizing our reasons for doing so.

Let us consider what kind of information is usually encoded in filesystem operations.
A usual set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$---create the object $\valv$ at node $n$,
$edit(n,\valv)$---modify the object at $n$ to become $\valv$,
and $remove(n)$---delete the object at node $n$; 
where $n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
edits that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors when combined with other commands are quite different.
Indeed, Bill Zissimopoulos' work 
demonstrated\footnote{Bill Zissimopoulos, personal communication, August 2015.}
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to encode information about the
output \emph{value} of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

Notice, however, that the commands listed above also encode some information about 
their input, the state of the filesystem
before the command is applied. In particular, $create(n,\valv)$ requires that there are no files
or directories at $n$, while $edit(n,\valv)$ and $remove(n)$ require the opposite.
This creates an arbitrary asymmetry where
there is now more information encoded into commands about their output than about their input.
As, based on the above, we expect that encoding more information in the commands
results in a model with greater predictive powers,
and in order to resolve this asymmetry, 
we propose a set of commands that encode
the type of the original state of $\FS(n)$ as well.
(Some real-life filesystem commands like $rmdir$ do this already.)

We therefore have four pieces of information related to each command:
its input type, its output type, the node to which it is applied, and finally
the value that it stores at the node.
We will denote commands using quadruples that list these,
and use $\cchard$, $\ccharf$ or $\ccharb$ to represent the different
input and output types. For example, $\cbdaa{n}{\vald}$ 
will represent $mkdir(n,\vald)$.
Alternatively, we use a triple of the two types and a node if 
there is only a single suitable output value for the given type (e.g. $\empt$).
That is, $\cdba{n}$ will represent $rmdir(n)$.

\medskip

A command can only succeed if the original value at node $n$ has a type that matches
the input type of the command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
{\em breaks} the filesystem. Broken filesystems are considered equal
(but not equal to any working filesystem), and we note them by $\FS=\fsbroken$.

@@

We note that a command or a sequence of commands is applied to a filesystem
by prefixing the command or sequence to it, for example: $\cbrk\FS$, $\cbdaa{n}{\vald}\FS$, 
or simply $S\FS$ if $S$ is a sequence of commands.

@@

The exact behavior of the commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
All other commands have the generic form $\cxynv$ where
$X$ and $Y$ are types in $\typeset$,
$n\in\setn$ and $\valvy$ is a value of the appropriate type from $\setvx{Y}$.
\[ \cxynv\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/n] \mbox{~violates the tree property}\\
   \FS[V/n] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

@@

\medskip

@@

For reasons also listed in \cite{NREC}, in this paper we will not consider
a $move$ or $rename$ command. Regarding the theoretical reasoning we aim to follow,
this turns out to be useful because these would be the only commands that affect
filesystems at two nodes at once, therefore describing 
the dependencies affecting these commands
would call for a more complicated model.
From a pragmatic perspective, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by pre- and post-processing changes in the filesystem to
discover them, which (especially when a rename is combined with changes to the content)
is a non-trivial problem in itself.

>>>>

Next we define commands on filesystems.
As described above, we aimed to select a set of commands
that captures as much information as possible
about the operations, and is also symmetric.
We start by summarizing our reasons for doing so.

Let us consider what kind of information is usually encoded in filesystem operations.
A usual set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$---create the value $\valv$ at node $n$,
$edit(n,\valv)$---modify the value at $n$ to become $\valv$,
and $remove(n)$---delete the value at node $n$; 
where $n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
$edit$s that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors are quite different when combined with other commands.
Indeed, Bill Zissimopoulos' work 
demonstrated\footnote{Bill Zissimopoulos, personal communication, August 2015.}
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to consider the
exact output \emph{value} of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

Notice, however, that the commands listed above also encode some information about 
their input, the state of the filesystem
before the command is applied. In particular, $create(n,\valv)$ requires that there are no files
or directories at $n$, while $edit(n,\valv)$ and $remove(n)$ require the opposite.
This creates an arbitrary asymmetry where
there is now more information available about their output than about their input.
As, based on the above, we expect that encoding more information in the commands
results in a model with greater predictive powers,
and in order to resolve this asymmetry, 
we propose a set of commands that encode
the type of the original state of $\FS(n)$ as well.
(Some real-life filesystem commands like $rmdir$ do this already.)

We therefore have four pieces of information related to each command:
its input type, its output type, the node to which it is applied, and finally
the value that it stores at the node.
We will denote commands using quadruples that list these
in a the following format:
\[ \cxynv, \]
where $X$ is the input type, $Y$ is the output type, $n$ is the node,
and $\valvy$ is a value of the appropriate type from $\setvx{Y}$.
We use $\cchard$, $\ccharf$ or $\ccharb$ to represent the different
types. For example, $\cbdaa{n}{\vald}$ 
will represent $mkdir(n,\vald)$.
Alternatively, we use a triple of the two types and a node if 
there is only a single suitable output value for the given type (e.g. $\empt$).
That is, $\cdba{n}$ will represent $rmdir(n)$.

\bigskip

\noindent
A command can only succeed if the original value at node $n$ has a type that matches
the input type of the command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
{\em breaks} the filesystem. Broken filesystems are considered equal
(but not equal to any working filesystem), and we note them by $\FS=\fsbroken$.

@@

We note that a command or a sequence of commands is applied to a filesystem
by prefixing the command or sequence to it, for example: $\cbrk\FS$, $\cbdaa{n}{\vald}\FS$, 
or simply $S\FS$ if $S$ is a sequence of commands. % TODO Add space between command and FS

@@

The exact behavior of the commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
\[ \cxynv\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/n] \mbox{~violates the tree property}\\
   \FS[V/n] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

@@

@@

\bigskip

\noindent
For reasons also listed in \cite{NREC}, in this paper we will not consider
a $move$ or $rename$ command. Regarding the theoretical reasoning we aim to follow,
this turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for $move$ would call for a more complicated model.
From a pragmatic perspective, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by pre- and post-processing changes in the filesystem to
discover them, which can enhance usability, but which
(especially when a rename is combined with changes to the content)
is a non-trivial problem in itself.

### 2015-12-23 2015/def-filesystem.tex
Comparison proof and text improvements

To model this, in $\setv$ we distinguish between 
file contents ($\setf$) and directory contents ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree property can then be defined as
\begin{align*}
\forall n\in\setn: &\FS(n) \neq \empt \\ % TODO alignment problem
&\quad\wedge \parent(n) \neq \topnode \\
&\Rightarrow \FS(\parent(n)) \in \setd 
\end{align*}
This means that as we move down from the root of a tree of nodes,
the types of values we encounter in the filesystem can only change according to the
transition diagram in \cref{fig_transition}, from directories ($\cchard$) to files ($\ccharf$)
to empty nodes ($\ccharb$).

@@

A simplification we will assume in this paper is regarding directories.
As Bill Zissimopoulos pointed 
out,\footnote{Bill Zissimopoulos, personal communication, August 2015.} 
we often do not want to consider metadata stored in
directories (e.g. permission settings) during synchronization,
as these are not generally understood by users,
and, if needed, conflict resolution on these settings can be easily automated.
In other words, the contents of directories are all equal,
which can be modelled by assuming that $|\setd|=1$.

>>>>

To model this, in $\setv$ we distinguish between 
file values ($\setf$) and directory values ($\setd$), that is,
if $\setb = \{\empt\}$ then:
\[ \setv = \setb \cup \setf \cup \setd \]
The tree property can then be defined as
\begin{align*}
\forall n\in\setn: &\FS(n) \neq \empt \\ % TODO alignment problem
&\quad\wedge \parent(n) \neq \topnode \\
&\Rightarrow \FS(\parent(n)) \in \setd 
\end{align*}
This means that as we move down from the root of a tree of nodes,
the types of values we encounter in the filesystem can only change according to the
transition diagram in \cref{fig_transition}, from directories ($\cchard$) to files ($\ccharf$)
to empty nodes ($\ccharb$).

@@

A simplification we will assume in this paper is regarding directories.
As Bill Zissimopoulos pointed 
out,\footnote{Bill Zissimopoulos, personal communication, August 2015.} 
we often do not want to consider metadata stored in
directories (e.g. permission settings) during synchronization,
as these are not generally understood well by users,
and, if needed, conflict resolution on these settings can be easily automated.
In other words, the contents of directories are all equal,
which can be modelled by assuming that $|\setd|=1$.

### 2015-12-23 2015/intro.tex
More improvements and references
Comparison proof and text improvements

A central problem of command-based synchronizers during both update detection and reconciliation
is ordering.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied
to the filesystems
without causing an error 
(for example, a directory needs to be created first before a file can be created under it).
Similarly, during reconciliation when we consider all the updates that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands,
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role.
Other research also shows the essential nature of ordering:
in IceCube multiple orders are tested to find an acceptable one \cite{KRSD},
or in Bayou reconciling updates happens by redoing them in an acceptable order \cite{TTPDSH}.

In \cref{sec_def} we define a model of filesystems and a set of commands
that we will use to describe updates and modifications.
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
\Cref{sec_update} describes update detection, and it is
in \cref{ordering} that we describe how commands can be ordered
and prove that under some conditions all valid orders
have the same effect.

Finally, \cref{sec_rec} defines our reconciliation 
and conflict detection algorithms.
We then proceed to prove that the output of reconciliation
is sound in the sense that 
it can be applied to the replicas without causing an error,
and that the described algorithm provides a complete solution
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
the concept of \emph{conditional operation} of sequences of commands,
and show that it has a number of highly convenient properties.

>>>>

A central problem of command-based synchronizers during both update detection and reconciliation
is ordering.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied
to the filesystems without causing an error 
(for example, a directory needs to be created before a file can be created under it).
Similarly, during reconciliation when we consider all the updates that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands,
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role.
Other research also shows the essential nature of ordering:
in IceCube multiple orders are tested to find an acceptable one \cite{KRSD},
or in Bayou reconciling updates happens by redoing them in an acceptable order \cite{TTPDSH}.

In \cref{sec_def} we define a model of filesystems and a set of commands
that we will use to describe updates and modifications.
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
\Cref{sec_update} describes update detection, and it is
in \cref{ordering} that we describe how commands can be ordered
and prove that under some simple conditions all valid orders
have the same effect.

\Cref{sec_rec} defines our reconciliation 
and conflict detection algorithms.
We then proceed to prove that the output of reconciliation
is sound in the sense that 
it can be applied to the replicas without causing an error,
and it is complete
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
the concept of \emph{conditional operation} of sequences of commands,
and show that it has a number of highly convenient properties.
We finish by comparing our results to other research
and outlining directions for further work
on the introduced algebraic system.

### 2015-12-23 2015/ordering.tex
Comparison proof and text improvements

\begin{mylem}\label{connected_changes}
Given a set of commands that
does not contain assertion commands,
and which can be applied to a filesystem in some order without breaking it,
and which contains commands on $n$ and $\nn$ where $\nn\descendant n$
but $\nn\neq\parent(n)$, then the set must also contain a command
on each node between $n$ and $\nn$.
\end{mylem}
\begin{proof}
We prove that under the given conditions, the set must contain a command on $\parent(n)$.
Then, by repapplying this result, we how that the set must contain commands on every
ancestor of $n$ up to $\nn$.

@@

Let $A$ be a sequence that satisfies the conditions;
we therefore know that it contains a command $\cxynv$ on $n$
and another command $\czwnnv$ on $\nn$.
We use an indirect proof and assume that there are no commands on $\parent(n)$ in $A$.
Next, we create a new sequence $A'\equiv A$ in which $\cxynv$ and $\czwnnv$ are next to each other.
If they are already next to each other in $A$, there is nothing to do.
Otherwise, consider the command next to $\cxynv$ in the direction where $\czwnnv$ is.
Let this command be $\cqrmv$.
If $\czwnnv$ is to the right, then $A$ looks like the following:
\[ A = \cdots\cc\cxynv\cc\cqrmv\cc\cdots\cc\czwnnv\cc\cdots \]
If $n\unrel m$, swap $\cxynv$ and $\cqrmv$. Based on \cref{ax_separate_commute} we know that the new
sequence is equivalent to $A$.
Otherwise, we know $m\neq\parent(n)$ as there are no commands on $\parent(n)$, and so
from \cref{ax_distantrel_breaks} we get $A\equiv\cbrk$ which contradicts our assumptions.
(Note that $A$ cannot contains assertion commands.)
By repeating this step we can therefore convert $A$ into $A'$ where $\cxynv$ and $\czwnnv$ are neighboring commands.
However, then \cref{ax_distantrel_breaks} applies to the sequence and therefore $A\equiv A'\equiv\cbrk$ which
is again a contradiction.
\end{proof}

@@

Let $U$ be a simple set of commands
that we know can be applied to at least one filesystem in some order without breaking it.
Then we can split $U$ into a maximum number of disjunct subsets
(components) so that the nodes of two commands from two different subsets are incomparable.
These will simply be the components of the subset of the node forest that contains the nodes affected by the set.

From \cref{ax_separate_commute} we know that commands from different 
components can be applied in any order.
Therefore once we have ordered the commands inside a component, we generate
all permutations of the components themselves, including the ones
in which commands from separate components mix and overlap
(but where the possible orders determined for commands inside the components are respected).

@@

In each component that is larger than one, 
we know that all nodes are related, and from \cref{connected_changes}
we also know that it must contain commands on all nodes in between.
Therefore the nodes
form a rooted ordered tree,
and the pairs of commands connected by the edges in the tree 
must be construction or destruction pairs.
(Otherwise, there would be no ordering of the component that is not equivalent to $\cbrk$
and via \cref{ax_separate_commute} there would be no ordering of the whole set, either.)

To order the commands in the component,
select the command with the topmost node, so that there would be no command
on its parent node (if there is one). 
If the command removes content at this node (it is of the form $\caa{X}{\ccharb}$), 
then it must follow all other commands in the component, and
all parent--child pairs in the whole component must be destruction pairs.
Otherwise, the command with the topmost node needs to precede the other commands,
and the component must be made of construction pairs.
These follow from the patterns of construction and destruction pairs, which
at no point can mix without causing any sequence to break all filesystems.

@@

As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.
An important property of simple sequences is that
all of their valid orderings are equivalent,
and all equivalent sequences are mere reorderings:

\begin{mylem}\label{simple_reorder_equiv}
For any simple sequence $S\neq\cbrk$,
$\forall T: T\equiv S \Leftrightarrow T\in\orderset{S}$.
\end{mylem}
\begin{proof}
First, as $S$ only has at most one command on each node,
any sequence in $\orderset{S}$ either breaks a filesystem,
or it leaves it in the same state as $S$.

@@

Therefore if $S\nequiv\cbrk$ then $S\in\orderset{S}$,
and $S$ cannot equivalent to any sequence of the same commands
not in $\orderset{S}$.
\end{proof}

>>>>

\begin{mylem}\label{connected_changes}
Given a set of commands that
does not contain assertion commands,
and which can be applied to a filesystem in some order without breaking it,
and which contains commands on $n$ and $\nn$ where $\nn\descendant n$
but $\nn\neq\parent(n)$, then the set must also contain a command
on each node between $n$ and $\nn$.
\end{mylem}
\begin{proof}
We prove that under the given conditions, the set must contain a command on $\parent(n)$.
Then, by repapplying this result, we know that the set must contain commands on every
ancestor of $n$ up to $\nn$.

@@

Let $A$ be a sequence that satisfies the conditions;
we therefore know that it contains a command $\cxynv$ on $n$
and another command $\czwnnv$ on $\nn$.
We use an indirect proof and assume that there are no commands on $\parent(n)$ in $A$.
Next, we create a new sequence $A'\equiv A$ in which $\cxynv$ and $\czwnnv$ are next to each other.
If they are already next to each other in $A$, there is nothing to do.
Otherwise, consider the command next to $\cxynv$ in the direction where $\czwnnv$ is.
Let this command be $\cqrmv$.
If $\czwnnv$ is to the right, then $A$ looks like the following:
\[ A = \cdots\cc\cxynv\cc\cqrmv\cc\cdots\cc\czwnnv\cc\cdots \]
If $n\unrel m$, then swap $\cxynv$ and $\cqrmv$. Based on \cref{ax_separate_commute} we know that the new
sequence is equivalent to $A$.
Otherwise, we know $m\neq\parent(n)$ as there are no commands on $\parent(n)$, and so
from \cref{ax_distantrel_breaks} we get $A\equiv\cbrk$ which contradicts our assumptions.
(Note that $A$ cannot contains assertion commands.)
By repeating this step we can therefore convert $A$ into $A'$ where $\cxynv$ and $\czwnnv$ are neighboring commands.
However, then \cref{ax_distantrel_breaks} applies to the sequence and therefore $A\equiv A'\equiv\cbrk$ which
is again a contradiction.
\end{proof}

@@

Let $U$ be a simple set of commands
that we know can be applied to at least one filesystem in some order without breaking it.
Then we can split $U$ into a maximum number of disjunct subsets
(components) so that the nodes of two commands from two different subsets are incomparable.
These will simply be the components of the subset of the node forest that contains 
the nodes affected by $U$.

From \cref{ax_separate_commute} we know that commands from different 
components can be applied in any order.
Therefore once we have ordered the commands inside the components, we generate
all permutations of the components themselves, including the ones
in which commands from separate components mix and overlap
(but where the possible orders determined for commands inside the components are respected).

@@

In each component that is larger than one, 
we know that all nodes are related, and from \cref{connected_changes}
we also know that it must contain commands on nodes in between the related nodes.
Therefore the nodes
form a rooted ordered tree,
and the pairs of commands connected by the edges in the tree 
must be construction or destruction pairs.
(Otherwise, there would be no ordering of the component that is not equivalent to $\cbrk$
and via \cref{ax_separate_commute} there would be no ordering of the whole set, either.)

To order the commands in the component,
select the command with the topmost node, so that there would be no command
on its parent node (if there is one). 
If the command removes content at this node (it is of the form $\caa{X}{\ccharb}$), 
then it must follow all other commands in the component, and
all parent--child pairs in the whole component must be destruction pairs.
Otherwise, the command with the topmost node needs to precede the other commands,
and the component must be made of construction pairs.
These follow from the patterns of construction and destruction pairs, which
at no point can mix without causing any ordering to break all filesystems.

@@

As a specific case, we can regard a simple sequence of commands $S$
as a set and determine its $\orderset{S}$.
An important property of simple sequences is that
all of their valid orderings are equivalent,
and all equivalent sequences are mere reorderings:

\begin{mylem}\label{simple_reorder_equiv}
For any simple sequence $S\nequiv\cbrk$,
$\forall T: T\equiv S \Leftrightarrow T\in\orderset{S}$.
\end{mylem}
\begin{proof}
First, as $S$ only has at most one command on each node,
any sequence in $\orderset{S}$ either breaks a filesystem,
or it leaves it in the same state as $S$.

@@

Therefore if $S\nequiv\cbrk$ then $S\in\orderset{S}$,
and $S$ cannot be equivalent to any sequence of the same commands
not in $\orderset{S}$.
\end{proof}
